---
title: "Exam 2"
subtitle: <center> <h1>Take-Home Data Analysis</h1> </center>
author: <center> Jake Hiltscher <center>
output: html_document
---

<style type="text/css">
h1.title {
  font-size: 40px;
  text-align: center;
}
</style>


```{r setup, include=FALSE}
# load any necessary packages here
library(tidyverse)
library(ggfortify)  # plot lm objects using ggplot instead of base R
library(car)  # needed for added-variable plots and dfbetas and dffits
library(corrplot)  # colored correlation matrix
library(gridExtra)  # NEW PACKAGE for putting multiple ggplots in one plot
library(bestglm)  # for stepwise methods
library(glmnet)  # for ridge, lasso, and elastic net
set.seed(12345)
```

## Exam Instructions

Use this .rmd file to perform your analysis. You will answer questions about this data set through a Canvas Quiz. You will then submit your completed .rmd and .html file at the end of the Canvas quiz. Your code should be organized, properly formatted, and you should only print out relevant items (e.g. do not print out the entire data set for me to see, and do not use the `View()` function).

## Data and Description

Researchers collected data on 414 homes in China. The goal is to use rather unique explanatory variables to predict house price (measured in cost (US Dollars) per unit area). The following are predictor variables:

Variable | Description
-------- | -------------------------
age      | Age of the house (number of years since being built) in years
stores   | Number of convenience stores within a 500 meter circle around the house
metro    | Distance from the house to the nearest metro station (in meters) 
trails   | Distance from the house to the nearest walking trail (in meters) 

Download the RealEstate.txt file from Canvas (Files -> DATA SETS), and put it in the same folder as this R Markdown file.

### Complete your exploratory data analysis (EDA) in this section. You may use multiple code chunks, if you wish, to organize your code.

```{r}
real_estate <- read_table("RealEstate.txt")
real_estate <- as.data.frame(real_estate)
summary(real_estate)
```

### Fit an inital model, and check the "(A) no influential points" assumption in this section. You may use multiple code chunks, if you wish, to organize your code.

```{r}
real_estate_lm <- lm(price ~ ., data = real_estate)
summary(real_estate_lm)
pairs(real_estate)

corrplot(cor(real_estate))

avPlots(real_estate_lm)

# age -------------------------------------------------------------------
# calculate the DFBETAS for age
real_estate$dfbetas_age <- as.vector(dfbetas(real_estate_lm)[, "age"])
# plot the DFBETAS against the observation number
dfbetas_plot_age <- ggplot(data = real_estate) + 
  geom_point(mapping = aes(x = as.numeric(rownames(real_estate)), 
                           y = abs(dfbetas_age))) +
  ylab("Abs(DFBETAS) for Age") +
  xlab("Observation Number") +
  # for n > 30
  geom_hline(mapping = aes(yintercept = 2 / sqrt(length(dfbetas_age))),
             color = "red", 
             linetype = "dashed") + 
  # for n <= 30
  geom_hline(mapping = aes(yintercept = 1),
             color = "red", 
             linetype = "dashed") +
  theme(aspect.ratio = 1)
# stores -------------------------------------------------------------------
# calculate the DFBETAS for stores
real_estate$dfbetas_stores <- as.vector(dfbetas(real_estate_lm)[, "stores"])
# plot the DFBETAS against the observation number
dfbetas_plot_stores <- ggplot(data = real_estate) + 
  geom_point(mapping = aes(x = as.numeric(rownames(real_estate)), 
                           y = abs(dfbetas_stores))) +
  ylab("Abs(DFBETAS) for Stores") +
  xlab("Observation Number") +
  # for n > 30
  geom_hline(mapping = aes(yintercept = 2 / sqrt(length(dfbetas_stores))),
             color = "red", 
             linetype = "dashed") + 
  # for n <= 30
  geom_hline(mapping = aes(yintercept = 1),
             color = "red", 
             linetype = "dashed") +
  theme(aspect.ratio = 1)
# metro -------------------------------------------------------------------
# calculate the DFBETAS for metro
real_estate$dfbetas_metro <- as.vector(dfbetas(real_estate_lm)[, "metro"])
# plot the DFBETAS against the observation number
dfbetas_plot_metro <- ggplot(data = real_estate) + 
  geom_point(mapping = aes(x = as.numeric(rownames(real_estate)), 
                           y = abs(dfbetas_metro))) +
  ylab("Abs(DFBETAS) for Metro") +
  xlab("Observation Number") +
  # for n > 30
  geom_hline(mapping = aes(yintercept = 2 / sqrt(length(dfbetas_metro))),
             color = "red", 
             linetype = "dashed") + 
  # for n <= 30
  geom_hline(mapping = aes(yintercept = 1),
             color = "red", 
             linetype = "dashed") +
  theme(aspect.ratio = 1)
# trails -------------------------------------------------------------------
# calculate the DFBETAS for trails
real_estate$dfbetas_trails <- as.vector(dfbetas(real_estate_lm)[, "trails"])
# plot the DFBETAS against the observation number
dfbetas_plot_trails <- ggplot(data = real_estate) + 
  geom_point(mapping = aes(x = as.numeric(rownames(real_estate)), 
                           y = abs(dfbetas_trails))) +
  ylab("Abs(DFBETAS) for Trails") +
  xlab("Observation Number") +
  # for n > 30
  geom_hline(mapping = aes(yintercept = 2 / sqrt(length(dfbetas_trails))),
             color = "red", 
             linetype = "dashed") + 
  # for n <= 30
  geom_hline(mapping = aes(yintercept = 1),
             color = "red", 
             linetype = "dashed") +
  theme(aspect.ratio = 1)
grid.arrange(dfbetas_plot_age,
             dfbetas_plot_stores,
             dfbetas_plot_metro,
             dfbetas_plot_trails,
             nrow = 2)

```

### Fit a model on a new data set where the one influential point is removed, and check the "no multicollinearity" assumption, in this section. You may use multiple code chunks, if you wish, to organize your code.

```{r}
# Removing influential point (Observation 271)
real_estate <- read_table("RealEstate.txt")
real_estate <- as.data.frame(real_estate)
new_real_estate <- real_estate[-271, ]

new_real_estate_lm <- lm(price ~ ., data = new_real_estate)
summary(new_real_estate_lm)
vif(new_real_estate_lm)

```

### Fit another model (on the new data set where the one influential point is removed) with the age variable square rooted (do not use the original age variable) in this section. 

```{r}
sqrt_real_estate_lm <- lm(price ~ sqrt(age) +.-age,data = new_real_estate)
summary(sqrt_real_estate_lm)

sqrt_real_estate <- new_real_estate
sqrt_real_estate$age_sqrt <- sqrt(sqrt_real_estate$age)

```

### Perform variable selection in this section. You may use multiple code chunks, if you wish, to organize your code.

```{r}
real_estate <- sqrt_real_estate[ , c("age_sqrt", "stores", "metro", "trails", "price")]
# Backwards selection
best_subsets_bic <- bestglm(real_estate,
                            IC = "BIC",
                            method = "exhaustive")

# view variables included in the top 10 models
best_subsets_bic$BestModels

# view a summary of the "best" model
summary(best_subsets_bic$BestModel)

# PMSE selection
best_subsets_pmse <- bestglm(real_estate,
                             IC ="CV")

# view variables included in the top 10 models
best_subsets_pmse

# view a summary of the "best" model
summary(best_subsets_pmse)
```

### Fit a model with any variables removed that you deem unnecesary for the model in this section.

```{r}
screen_real_estate <- sqrt_real_estate[ , c("age_sqrt", "stores", "metro", "price")]
summary(screen_real_estate)

screen_real_estate_lm <- lm(price ~., data = screen_real_estate)
summary(screen_real_estate_lm)
```

### Determine if any interaction(s) are needed for this model in this section. You may use multiple code chunks, if you wish, to organize your code.

```{r}
screen_real_estate_lm <- lm(price ~.+age_sqrt*stores+age_sqrt*metro+stores*metro, data = screen_real_estate)
summary(screen_real_estate_lm)
anova(screen_real_estate_lm)
screen_real_estate_lm <- lm(price ~.+age_sqrt*metro+stores*metro, data = screen_real_estate)
```
Two interactions are needed - Age:metro and stores:metro.

### Complete statistical inference based on the best linear model you chose in this section. You may use multiple code chunks, if you wish, to organize your code.

```{r}
# <your code here>
pairs(screen_real_estate)
corrplot(cor(screen_real_estate))
```
Through our analyses, the price of a house is effected by the number of metros and stores in an area, as well as the age of the house. There is also an interaction between number of metros and stores in an area. This is analysis is almost quantitatively comparing rural areas to metropolitan areas. Age also has an effect on the price of a house.

```{r}
predict(screen_real_estate_lm, 
        newdata = data.frame(age_sqrt = 20, stores = 3, metro = 500, trails = 6000),
        level = 0.95,
        interval = "confidence")
```